---
title: ClimateBC and ClimateNA
subtitle: Climate normals and projections
date: Nov 8, 2024
---

Websites:

- https://climatebc.ca/
- https://climatena.ca/

The new ClimateNAr library enables the downloading of up to 800m resolution climate normals and projections for BC and western NA. In this vignette, we will illustrate the use of the `ClimateNAr` package along with the `terra` package for reading and writing rasters. The `ClimateNAr` package needs to be downloaded and installed manually following the instructions:

1. Go to https://register.climatena.ca/ and register (it's free)
2. Download the ClimateBC/NA R package and documentation
3. Start R, go to the Packages menu, and select "Install package(s) from local file"
4. Once installed, restart R and load the library: library(ClimateNAr)

```{r}
#| message: false
#| warning: false
library(terra)
library(ClimateNAr)
```

In this short vignette, we will illustrate the use of the `ClimateNAr` package to download normal and projected data for an area of interest defined by a DEM raster. The steps include:

1. Create a boundary raster in tif format using a 900m resolution; make sure that the tif is in lat/long ("EPSG:4326").
3. Use the climateNAr function to download the files. The function has 4 parameters (make sure to read the package documentation for more details):

  - inputFile: the name of the tif dem raster
  - varList: the climate variables of interest
  - periodList: the names of the normal and projected scenarios
  - outDir: the name of the output directory (this will be created by the function)

### Read the DEM raster (needed for clipping)

```{r}
#| message: false
#| warning: false
r = rast('dem900m.tif')
plot(r)
```

### Enter parameters for climateNAr function

```{r}
#| message: false
#| warning: false
inputFile = 'dem900m.tif'
varList = c('MAT','MAP','DD5','CMI','RH')
periodList = c('Normal_1961_1990.nrm','Year_1902.ann','8GCMs_ensemble_ssp245_2041-2070.gcm')
outDir = 'download/'
if (!dir.exists(outDir)) dir.create(outDir)
```

### Download the climate data and plot

```{r}
#| message: false
#| warning: false
test <- climateNAr(inputFile, periodList, varList, outDir)
print(test)
plot(test)
```

### Clip the climate data to the DEM boundary

```{r}
#| message: false
#| warning: false
crs(test) <- "EPSG:4326" # set CRS to match DEM
test_mask <- mask(test, r) 
```

### Plot the downloaded raster

```{r}
#| message: false
#| warning: false
plot(test_mask)
```
